Secure Composition of Robust and Optimising Compilers

### 2023-06-22 popl/main.tex
Add figure

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
  \mi{(Pre\text{--}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \end{aligned}
\end{gather*}

@@

\MKin{sketch proof with backtranslation and the x-lang rels}

@@

>>>>

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{--}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}

@@

@@

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[state/.style={minimum height=0.6cm}]
      % relative horizontal/vertical distance between states
      \pgfmathsetmacro{\hdist}{0.95}
      \pgfmathsetmacro{\vdist}{1.35}
      \pgfmathsetmacro{\halfvdist}{0.725}

      % row of src states
      \node[state] (srcempty) {$\src{\emptyset}$};
      \foreach \s [remember=\s as \cur (initially empty)] in {1,w_1,p,w_2,2} {
        \node[state,right=\hdist of src\cur] (src\s) {$\src{\Omega_{\s}}$};
      }
      % row of trg states
      \node[state,below=\vdist of srcempty] (trgempty) {$\trg{\emptyset}$};
      \foreach \s in {1,w_1,p,w_2,2} {
        \node[state,below=\vdist of src\s] (trg\s) {$\trg{\Omega_{\s}}$};
      }
      %% illustrations
        % backtrans wrapper 1
        \draw[thick,loosely dotted,Peach!50,rounded corners] (src1.north east) -- (srcw\string_1.north east)
          -- (trgw\string_1.south east) -- (trg1.south west) -- (src1.north west) -- cycle;
        % backtrans wrapper 2
        \draw[thick,loosely dotted,Peach!50,rounded corners] (srcp.north east) -- (srcw\string_2.north east)
          -- (trgw\string_2.south east) -- (trgp.south west) -- (srcp.north west) -- cycle;
        % compiler correctness
        \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(srcw\string_1.north east)+(0,0.05)$) -- ($(srcp.north east)+(0,0.05)$)
          -- ($(trgw\string_2.north east)+(0.05,0)$) -- ($(trgw\string_2.south east)+(0.05,-0.05)$)
          -- ($(trg1.south west)+(-0.05,-0.05)$) -- ($(trg1.north west)+(-0.05,0)$) -- ($(srcw\string_1.north west)+(0,0.05)$) -- cycle;
      % state relations
      \path (srcempty) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_\emptyset$} (trgempty)
        (src1) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trg1)
        (srcp) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgp)
        (src2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_2}$} (trg2)
        (srcw\string_1) edge[draw=gray!25] node[pos=0.5,sloped,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trgw\string_1)
        (srcw\string_2) edge[draw=gray!25] node[pos=0.5,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (trgw\string_2)
        % diagonals
        (srcw\string_1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\approx_{\delta_{w_1}}$} (trg1)
        (src1) edge[draw=gray!25] node[pos=0.16,sloped,rotate=180,fill=white] {\scriptsize$\multimap_{\delta_1}$} (trgw\string_1)
        (srcp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\approx_{\delta_p}$} (trgw\string_2)
        (trgp) edge[draw=gray!25] node[pos=0.2,sloped,fill=white] {\scriptsize$\multimap_{\delta_{w_2}}$} (srcw\string_2)
        ;
      %\drawpolygon src1,srcw\string_1,trgw\string_1,trg1;
      %\drawpolygon srcp,srcw\string_2,trgw\string_2,trgp;
      %\node[font=\tiny,align=center,above=0.2 of srcw1srcp] (wrapper) {Backtranslation\\Wrapper};
      %\path[->,draw] (wrapper) -- (srcw\string_1);
      %\path[->,draw] (wrapper) -- (srcp);
      % steps
      \path[color=\stlccol] (srcempty) edge[draw=none] node {\ $\xrightarrow{\trace[_1]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src1)
        (src1) edge[draw=none] node {\ $\xrightarrow{\trace[_c]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_1)
        (srcw\string_1) edge[draw=none] node {\ $\xrightarrow{\trace[_p]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (srcp)
        (srcp) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\trace[_r]}{}{\kern-3.5pt}_{\text{ctx}}$} (srcw\string_2)
        (srcw\string_2) edge[draw=none] node {\ $\xrightarrow{\trace[_2]}{}{\kern-3.5pt}_{\text{ctx}}^*$} (src2)
        ;
      \path[color=\ulccol] (trgempty) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_1]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg1)
        (trg1) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_1)
        (trgw\string_1) edge[draw=none] node[fill=white,inner sep=0,outer sep=0] {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgp)
        (trgp) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_p]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trgw\string_2)
        (trgw\string_2) edge[draw=none] node {\ $\xrightarrow{\phantom{\trace[_2]}}{}{\kern-3.5pt}_{\text{ctx}}^*$} (trg2)
        ;
      % legend
      \node[align=left,below right=0.3 and 0.3 of trgempty,font=\tiny] (legend) {%
        ${\src{\trace[_c]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Call\ ?\finalexpr_i})}$\\%
        ${\src{\trace[_r]}}\cong{\backt{\Ltms}{\Ltrg}(\trg{Ret\ !\finalexpr_o})}$
        };
      \draw[thick,loosely dotted,Peach!50,rounded corners] ($(legend.north east)+(0.5,-0.4)$) -- ($(legend.north east)+(1,-0.4)$);
      \node at ($(legend.north east)+(1.5,-0.4)$) (legendwrapper) {};
      \node[right of=legendwrapper] {\tiny Backtranslation Wrapper};
      \draw[thick,loosely dashed,Emerald!50,rounded corners] ($(legend.south east)+(0.5,0.4)$) -- ($(legend.south east)+(1,0.4)$);
      \node at ($(legend.south east)+(1.5,0.4)$) (legendcorrectness) {};
      \node[right of=legendcorrectness] {\tiny Compiler Correctness};
    \end{tikzpicture}
    \caption{Proof diagram for \Cref{thm:cca:rtp:tms} depicting the general structure of robust preservation proofs.}\label{fig:proofdiag:rtp}
  \end{center}
\end{figure}
